Predicting Probability of Participants' Vote of Conservative and Liberty in 2025 Canadian election 

### 2024-04-18 paper 2/paper(1).qmd
Add files via upload

The table details the point estimates for each predictor along with their standard errors and confidence intervals, providing insight into the precision of the estimates and the potential range within which the true value of the coefficients may lie. Notably, the intercept and coefficients for variables such as 'provincePrince Edward Island' and 'income_before_taxUpper Class' have both negative and positive values, indicating their varied influence on the outcome variable. The confidence intervals suggest the level of certainty about the estimates, where a narrower interval represents a more precise estimate. These results contribute to our understanding of the factors influencing voter behavior, guiding the development of targeted strategies for engagement and policy-making.

@@

From top to bottom, provinces are listed in order of descending support for the Liberal party. For example, Newfoundland and Labrador exhibit the highest point estimate, indicating a stronger inclination towards the Liberal party, whereas Prince Edward Island shows a significantly lower estimate of Liberal support. There is no dashed line at the 50% mark here, unlike in the referenced Trump support graph, because the graph aims to show the varying levels of Liberal support across provinces rather than a majority threshold.

The variability in the error bars suggests that certain provinces, such as Ontario and Newfoundland and Labrador, have more precise estimates of Liberal support, while provinces like Nova Scotia have wider confidence intervals, indicating less certainty in the estimated proportion. This graphic presentation allows for an at-a-glance comparison of Liberal party support among the provinces, highlighting regions where the party is either particularly strong or may need to focus more campaign efforts.
\newpage

@@

```
@fig-liberal-votes-by-education depicts the estimated proportions of liberal votes across different education levels. Each point on the graph represents the mean proportion of votes for the Liberal party within each educational category. The horizontal error bars signify the confidence intervals around these estimates, indicating the degree of uncertainty or the range within which the true proportion is likely to fall.

@@

This visual analysis suggests a relationship between education level and voting patterns, with the 'High Education' group showing the highest estimated support for the Liberal party. In contrast, the 'Low Education' group appears to have the lowest, potentially indicating differing political preferences based on education. These insights could guide campaign strategies to address the educational divide in political support.

@@

## Consideration of Potential Bias

@@

@@

```{r echo=FALSE, message=FALSE, warning=FALSE, results='hide'}

cons_predictions <- predict(step_cons, newdata = step_cons_counts, type = "response", interval = "confidence")
cons_CI <- data.frame(cons_predictions)

cons_CI$lwr <- cons_CI$cons_predictions - 1.96 * sqrt(cons_CI$cons_predictions * (1 - cons_CI$cons_predictions) / nrow(cons_CI))
cons_CI$upr <- cons_CI$cons_predictions + 1.96 * sqrt(cons_CI$cons_predictions * (1 - cons_CI$cons_predictions) / nrow(cons_CI))

cons_CI$observation <- 1:nrow(cons_CI)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, results='hide'}

step_lib_counts <- cleaned_survey_data %>%
  group_by(age, sex, province, education, religion, income_before_tax) %>%
  summarise(N_j=n(), .groups = 'drop')  # Use .groups = 'drop' to remove the grouping

```{r echo=FALSE, message=FALSE, warning=FALSE}
lib_predictions <- predict(step_lib, newdata = step_lib_counts, type = "response", interval = "confidence")
lib_CI <- data.frame(lib_predictions)

lib_CI$lwr <- lib_CI$lib_predictions - 1.96 * sqrt(lib_CI$lib_predictions * (1 - lib_CI$lib_predictions) / nrow(lib_CI))
lib_CI$upr <- lib_CI$lib_predictions + 1.96 * sqrt(lib_CI$lib_predictions * (1 - lib_CI$lib_predictions) / nrow(lib_CI))

lib_CI$observation <- 1:nrow(lib_CI)

```{r, echo=FALSE, message=FALSE, warning=FALSE}

ggplot(cons_CI, aes(x = cons_predictions)) +
  geom_density(alpha = 0.5, fill = "blue") +
  labs(x = "Predicted Probability", y = "Density", title = "Plot.9 The Density Plot for Predicted Probabilities for Conservative")
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}

ggplot(lib_CI, aes(x = lib_predictions, fill = "Liberal")) +
  geom_density(alpha = 0.5, fill = "green") +
  labs(x = "Predicted Probability", y = "Density", title = "Plot.10 The Density Plot for Predicted Probabilities for Liberal")
```

>>>>

The table details the point estimates for each predictor along with their standard errors and confidence intervals, providing insight into the precision of the estimates and the potential range within which the true value of the coefficients may lie. Notably, the intercept and coefficients for variables such as 'provincePrince Edward Island' and 'provinceOntario' have both negative and positive values, indicating their varied influence on the outcome variable. The confidence intervals suggest the level of certainty about the estimates, where a narrower interval represents a more precise estimate. These results contribute to our understanding of the factors influencing voter behavior, guiding the development of targeted strategies for engagement and policy-making.

@@

From top to bottom, provinces are listed in order of increasing support for the Liberal party. For example, Newfoundland and Labrador exhibit the highest point estimate (about 39%), indicating a stronger inclination towards the Liberal party, whereas Prince Edward Island shows a significantly lower estimate of Liberal support (about 0%). The graph aims to show the varying levels of Liberal support across provinces, the black dot on the graph indicates the point estimate of the proportion of voters within the corresponding province who are predicted to vote for Liberal party.

The variability in the error bars suggests that certain provinces, such as Ontario and Newfoundland and New Brunswick, have more precise estimates of Liberal support, while provinces like Nova Scotia have wider confidence intervals, indicating less certainty in the estimated proportion. This graphic presentation allows for an at-a-glance comparison of Liberal party support among the provinces, highlighting regions where the party is either particularly strong or may need to focus more campaign efforts.
\newpage

@@

```
@fig-liberal-votes-by-education depicts the estimated proportions of liberal votes across different education levels. Each point on the graph represents the mean proportion of votes for the Liberal party within each educational category. The horizontal error bars signify the confidence intervals around these estimates, indicating the degree of uncertainty or the range within which the true proportion is likely to fall. Moreover, the black dot notes the point estimate of the proportion of voters within the corresponding education category who are predicted to vote for liberal party.

@@

This visual analysis suggests a relationship between education level and voting patterns, with the 'High Education' group showing the highest estimated support for the Liberal party (about 33%). In contrast, the 'Low Education' group appears to have the lowest (about 23%), potentially indicating differing political preferences based on education. These insights could guide campaign strategies to address the educational divide in political support.

@@

## Consideration of Potential Bias and Limitations

@@

```{r echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
step_lib_counts <- cleaned_survey_data %>%
  group_by(age, sex, province, education, religion, income_before_tax) %>%
  summarise(N_j=n(), .groups = 'drop')

```{r echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
step_lib_counts <- cleaned_survey_data %>%
  group_by(age, sex, province, education, religion, income_before_tax) %>%
  summarise(N_j=n(), .groups = 'drop')

```{r echo=FALSE, message=FALSE, warning=FALSE, results='hide', include =FALSE}

lib_predictions <- predict(step_lib, newdata = step_lib_counts, type = "response")

lib_CI <- data.frame(
  estimate = lib_predictions,

  observation = seq_along(lib_predictions)
)

head(lib_CI)

```{r}
#| label: fig-liberal-votes-and-consrvative_votes
#| fig-cap: "predicted probabilities for the Conservative and Liberal parties"
#| echo: false
cons_probs <- predict(step_cons, newdata = cleaned_survey_data, type = "response")
lib_probs <- predict(step_lib, newdata = cleaned_survey_data, type = "response")

combined_predictions <- data.frame(
  PredictedProbability = c(cons_probs, lib_probs),
  Party = rep(c("Conservative", "Liberal"), each = length(cons_probs))
)

ggplot(combined_predictions, aes(x = PredictedProbability, fill = Party)) +
  geom_density(alpha = 0.5) +
  facet_wrap(~ Party, nrow = 2) +
  scale_fill_manual(values = c("red", "green")) +
  labs(x = "Predicted Probability", y = "Density") +
  theme_minimal() +
  theme(legend.position = "none")

@fig-liberal-votes-and-consrvative_votes provide density plots illustrating the distribution of predicted probabilities for the Conservative and Liberal parties, respectively. Both plots reveal a slight left skew, indicating a greater concentration of predictions at the lower probability range, nearer to 0.2. This skew suggests that while both parties are predicted to have a significant chance of receiving votes, there's a tendency for the model to assign a moderate likelihood rather than a strong certainty of voting outcomes.

Notably, the Liberal party's density plot reaches a higher peak, which points to a narrower, more defined range of predicted probabilities. This contrast implies that the model predicts voting probabilities for the Liberal party with more confidence, as the predictions are less spread out over a range of values compared to the Conservative party.

Additionally, the visible overlap between the two distributions signifies a degree of ambiguity or uncertainty in differentiating the predicted probabilities for the two parties. The shared area of the plots indicates that for a substantial portion of predictions, the likelihood of individuals voting for either party is comparably uncertain.

This graphical representation underscores the nuanced nature of electoral predictions, where despite clear tendencies, there remains a non-negligible degree of overlap and hence, predictive uncertainty. It is important to note that the analysis does not incorporate predictions for the NDP party, focusing solely on the Conservative and Liberal parties.

@@
